# Capital region

## Data

### Parishes in the capital region

```{r data-muni}

# Download DAGI (scale 1:10000)
  dangeo_get_data(ftp_folder = "landinddelinger/dagi/SHAPE",
                  zip_name   = "DAGIREF_SHAPE_UTM32-EUREF89.zip")

# Municipalities (get from DAGI)
  muni_link <- paste(loc_dir,
                        "DAGIREF_SHAPE_UTM32-EUREF89/ADM",
                        "KOMMUNE.shp",
                        sep = "/")
  dk_muni  <- read_sf(muni_link) %>%
    st_zm() %>% 
    st_transform(crs = "EPSG:25832") %>% 
    select(KOMKODE, KOMNAVN) %>% 
    rename(muni_id = KOMKODE,
           muni_name = KOMNAVN) %>% 
    group_by(muni_id, muni_name) %>%
    summarise(geometry = sf::st_union(geometry)) %>%
    ungroup() %>% 
    # Calculate the area of each municipalitie (km2) 
    mutate(area_km2 = as.numeric(units::set_units(st_area(.), km^2))) 
  
# Country
  dk_country <- st_union(dk_muni)
  
# Codes of the communes under study
  capital_region_muni <- c("København",
                           "Frederiksberg",
                           "Albertslund",
                           "Ballerup",
                           "Brøndby",
                           "Dragør",
                           "Gentofte",
                           "Gladsaxe",
                           "Glostrup",
                           "Herlev",
                           "Høje-Taastrup",
                           "Hvidovre",
                           "Ishøj",
                           "Lyngby-Taarbæk",
                           "Rødovre",
                           "Tårnby",
                           "Vallensbæk")

# Select municipalities under study  
 capital_muni <- dk_muni %>% filter(muni_name %in% capital_region_muni) 
  
 dk_country_crop <- st_crop(dk_country, capital_muni)
 
# Parishes polygons of Denmark, and select those in the study area 
  prsh_link <- paste(loc_dir,
                       "DAGIREF_SHAPE_UTM32-EUREF89/ADM",
                       "SOGN.shp",
                       sep = "/")
  dk_prsh  <- read_sf(prsh_link) %>%
    st_zm() %>% 
    st_transform(crs = "EPSG:25832")

# Select those where the centroid is in the capital region
  dk_prsh_cent <- st_centroid(dk_prsh)
  capital_prsh_cent <- st_intersection(dk_prsh_cent, capital_muni)

  capital_prsh <- dk_prsh %>% 
    # Get parishes in the capital region
    filter(SOGNEKODE %in% capital_prsh_cent$SOGNEKODE) %>% 
    # Combine several parish features geometries into one polygon
    group_by(SOGNEKODE, SOGNENAVN) %>% 
    summarise(geometry = st_union(geometry)) %>% 
    ungroup() %>% 
    # add area of the parish (in km2)
    mutate(prsh_area_km2 = as.numeric(units::set_units(st_area(.), km^2))) %>% 
    # Translate column names
    rename(prsh_id = SOGNEKODE,
           prsh_name = SOGNENAVN) 

# Contour of the capital region (merge the parishes in one polygon):
  capital_area <- capital_prsh %>%
    st_union() %>%
    st_sf() %>% 
    st_transform(crs = "EPSG:25832")
  
```

```{r fig-cap-reg-prsh}

ggplot() + 
  geom_sf(data = capital_prsh, 
          aes(geometry = geometry),
          fill = "grey", 
          color = "grey50", 
          size = 0.05) +
  geom_sf(data = capital_muni, 
          aes(geometry = geometry),
          fill = NA, 
          color = "black",
          size = 0.5) +
  theme_void() +
  annotation_scale(location = "br", text_cex = 1) +
  annotation_north_arrow(location = "br",
                         pad_x = unit(2, "cm"),
                         pad_y = unit(0.65, "cm"),
                         which_north = "true",
                         height = unit(0.5, "cm"),
                         width = unit(0.5, "cm"),
                         style = north_arrow_orienteering(text_col = "white",
                                                          text_size = 1))

```

### Population 

Population data from Denmark Statistics:

-   [KMSTA001: Population 1. January by parish, ancestry and National Church](https://www.statbank.dk/statbank5a/SelectVarVal/Define.asp?MainTable=KMSTA001&PLanguage=1&PXSId=0&wsid=cftree).

```{r aux-functions}

## Auxiliary functions for reading the data with the package *danstat*

# Loop by year for getting DST data 
  steps <- function(year){
    var_values <- list(id_region, id_ancestry, year)
    var_input <- purrr::map2(.x = var_codes,
                             .y = var_values,
                             .f = ~list(code = .x, values = .y))
    get_data(id_table, variables = var_input)
  }
 
# Function for removing punctuation, lowercase, stem, stopwords, and collapse strings
  rm_words <- function(x, stopwords) { x %>% 
      strsplit(" ", fixed = TRUE) %>% 
      lapply(tm::removePunctuation) %>% 
      lapply(tolower) %>% 
      lapply(SnowballC::wordStem) %>% 
      lapply(function(x) x[!x %in% stopwords]) %>% 
      vapply(function(x) paste(x , collapse = "_"), character(1))
  }
  
```

```{r pop-parish-KMSTA001}

## Read and clean table KMSTA001
 
# Table 
  id_table <- "KMSTA001"
  var_pop <- get_table_metadata(table_id = id_table, variables_only = TRUE)

# Codes for var_input
  var_codes <- c("SOGN", "HERKOMST", "Tid")
  
# Values for var_input
  # Region: parishes of the study area (i.e. capital_parish)
  id_region <- capital_prsh$prsh_id 
  # Ancestry
  id_ancestry <- NA
  # Quarters
  id_year <- var_pop$values[[4]]$id[1:13]   # Select 2008-2020
  
# Read data (n parallel) 
  plan(multisession, workers = 7)
  capital_prsh_ancestry_read <- id_year %>%
    future_map_dfr(steps)  
  plan("default")

# Clean data 
  capital_prsh_ancestry <- capital_prsh_ancestry_read %>% 
    # Translate column names into English
    rename(parish = SOGN,
           ancestry = HERKOMST,
           year = TID, 
           value = INDHOLD) %>%
    # Get parish codes, names, and municipality names
    separate(parish,
             c("prsh_id", "prsh_name", "muni_name"),
             sep = " ",
             extra = "drop") %>% 
    mutate(muni_name = gsub("\\(", "", muni_name)) %>% 
    # Make shorter names in ancestry
    mutate(ancestry = case_when(
      ancestry == "Persons of Danish origin" ~ "pop_dan",
      ancestry == "Immigrants from western countries" ~ "pop_mi_wst",
      ancestry == "Immigrants from non-western countries" ~ "pop_mi_nwst",
      ancestry == "Descendants from western countries" ~ "pop_de_wst",
      ancestry == "Descendants from non-western countries" ~ "pop_de_nwst"), 
      ancestry = factor(ancestry)) %>% 
    # Pivot (one row for peach parish and year)
    pivot_wider(names_from = ancestry, values_from = value) %>% 
    # Merge immigrants and their descendants (i.e. foreigners) 
    mutate(pop_frgn_wst = pop_mi_wst + pop_de_wst, 
           pop_frgn_nwst = pop_mi_nwst + pop_de_nwst) %>% 
    select(-c(pop_mi_wst, pop_de_wst, pop_mi_nwst, pop_de_nwst)) %>% 
    # Add column with total population
    mutate(pop_total = select(., starts_with("pop_")) %>% rowSums()) %>% 
    # Put NA when pop_* is 0
    mutate(across(starts_with("pop"), ~ifelse(.x == 0, NA, .x)))

  # Add the spatial information:

capital_prsh_ancestry_sf <- capital_prsh %>%
  select(prsh_id, prsh_area_km2) %>% 
  left_join(capital_prsh_ancestry, by = c("prsh_id")) %>% 
  # Population density
  mutate(across(starts_with("pop"), ~.x/prsh_area_km2, .names = "{.col}_km2"))

```

### Housing prices

```{r fig-cap-reg-prices, fig.height=10, fig.width=10, fig.cap = "Median 2020-adjusted house prices in the ordinary free trade by grid cells of 100m x 100m"}

# Load data
sum_runits_oft_prices <- readRDS("prsh_house_prices.rds")

# Plot
ggplot() +
  geom_sf(data = dk_country_crop,
          aes(geometry = geometry),
          fill = "grey") +
  geom_sf(data = sum_runits_oft_prices, 
          aes(geometry = geometry,
              fill = cut_number(median_2020_kDKK_m2,
                                n = 10,
                                ordered_result = TRUE,
                                dig.lab = 0)),
          color = NA,
          legend = "meh") +
  geom_sf(data = capital_muni,
          aes(geometry = geometry),
          fill = NA,
          color = "white",
          size = 0.05) +
  scale_fill_viridis_d(
    name = TeX("$\\overset{\\textbf{Percentiles}}{(kDKK/m^2)}$"),
    option = "magma",
    direction = -1) +
  theme_void() +
  guides(fill = guide_legend(reverse=T)) +
  labs(x = "",
       y = "") +
  facet_wrap( ~year, ncol = 3)

```

## Ternary plots

### Diagrams

Population structure by year, with all the data together in the same diagram. 

```{r fig-cap-reg-tern-pop-year, fig.cap = "Ternary plot by year"}

ggtern::ggtern(data = capital_prsh_ancestry_sf,
       aes(x = pop_dan,
           y = pop_frgn_wst,
           z = pop_frgn_nwst,
           colour = factor(year))) +
  scale_color_viridis_d(name = "Year") +
  geom_point() +
  ggtern::theme_rgbw() +
  ggtern::theme_hidetitles()

```

Separate plots by year and add the parishes in the legend. 

```{r fig-cap-reg-tern-pop-year-prsh, fig.height=20, fig.width=10, fig.cap = "Ternary plots by parish"}

ggtern::ggtern(data = capital_prsh_ancestry_sf,
       aes(x = pop_dan,
           y = pop_frgn_wst,
           z = pop_frgn_nwst,
           colour = prsh_name)) +
  geom_point() +
  facet_wrap( ~year, ncol = 3) +
  ggtern::theme_rgbw() +
  ggtern::theme_hidetitles() +
  theme(legend.position = "bottom") +
  guides(colour = guide_legend(title.position = "top"))

```

Links with house prices (zoom the figure to the parishes with median values). 

```{r fig-cap-reg-tern-pop-price-all, fig.cap = "Median housing prices and popupation distribution by parish (all data)"}

ggtern::ggtern(data = sum_runits_oft_prices,
               aes(x = pop_dan,
                   y = pop_frgn_wst,
                   z = pop_frgn_nwst,
                   colour = cut_number(median_2020_kDKK_m2,
                                       n = 10,
                                       dig.lab = 0))) +
  viridis::scale_color_viridis(name = "Percentiles\n[kDkk/m2]",
                               option = "turbo",
                               discrete = TRUE) +
  geom_point(size = 0.85) +
  ggtern::theme_rgbw() +
  ggtern::theme_hidetitles() +
  ggtern::theme_zoom_L(0.6) +
  guides(colour = guide_legend(reverse = TRUE,
                               override.aes = list(size = 3)))

```

```{r fig-cap-reg-tern-pop-price-year, fig.height=10, fig.width=10, fig.cap = "Median housing prices and popupation distribution by parish and year"}

ggtern::ggtern(data = sum_runits_oft_prices,
               aes(x = pop_dan,
                   y = pop_frgn_wst,
                   z = pop_frgn_nwst,
                   colour = cut_number(median_2020_kDKK_m2,
                                       n = 10,
                                       dig.lab = 0))) +
  facet_wrap( ~ year, ncol = 3)+
  viridis::scale_color_viridis(name = "Percentiles\n[kDkk/m2]",
                               option = "turbo",
                               discrete = TRUE) +
  geom_point(size = 0.9) +
  ggtern::theme_rgbw() +
  ggtern::theme_hidetitles() +
  ggtern::theme_zoom_L(0.6) +
  guides(colour = guide_legend(reverse = TRUE,
                               override.aes = list(size = 3)))

```

### Maps

Center over the compositional mean of 2019 (i.e. Danes ~ 80.3%; non-wst ~ 12.3; and wst ~ 7.4%).  

```{r fig-cap-reg-tern-map, fig.height=20, fig.width=20, fig.cap="Population distribution"}

f_tern_map <- function(df) {

# Make colours
tric <- tricolore::Tricolore(df,
                             p1 = "pop_dan",
                             p2 = "pop_frgn_wst",
                             p3 = "pop_frgn_nwst",
                             breaks = Inf,
                             show_data = FALSE,
                             center = c(0.803, 0.074, 0.123),
                             hue = 2/12,
                             lightness = 1,
                             chroma = 1
                             )

# legend
p_legend <- tric$key +
  labs(L = "% Danes",
       T = "% Wst",
       R = "% Non-wst") +
  theme(axis.title = element_text(size = 5, face = "bold"),
        axis.text  = element_text(size = 5))

# Add columns with colours
df <- df %>%
  mutate(pop_rgb = tric$rgb)

# Map
ggplot() +
  geom_sf(data = dk_country_crop,
          aes(geometry = geometry),
          fill = "grey") +
  geom_sf(data = df,
          aes(fill = pop_rgb, geometry = geometry),
          size = 0.05) +
  scale_fill_identity() +
  geom_sf(data = capital_muni,
          aes(geometry = geometry),
          fill = NA,
          color = "white",
          size = 0.5)+
  labs(title = df$year) +
  theme_void() +
  annotation_custom(
    ggtern::ggplotGrob(p_legend) ,
    xmin = 730000,
    xmax = 742000,
    ymin = 6178000,
    ymax = 6190000
  )

}

plts <- capital_prsh_ancestry_sf %>%
  group_split(year) %>%
  map(. , .f = f_tern_map)

wrap_plots(plts, ncol = 3)

```

Only for 2020.

```{r fig-cap-reg-tern-map-2020, fig.cap="Population distribution in 2020"}

plts[[13]]

```




